### 2021-06-09 paper-calculations.Rmd
some cleaning

---
title: "R Notebook"
output: html_notebook
---

Install required packages

```{r}
devtools::install_github("teunbrand/ggh4x")
```

Load packages

```{r}
library(readxl)
library(tidyverse)
library(xtable)
library(lubridate)
library(ggrepel)
library(ggh4x)
library(haven)

theme_set(theme_bw())
```

Selectivv data

```{r}
selectivv <- list()
selectivv[[1]] <- read_csv(file = "data/gig-table1-monthly-counts-stats.csv")
selectivv[[2]] <- read_csv(file = "data/gig-table2-halfyear-demo-stats.csv")
selectivv[[3]] <- read_csv(file = "data/gig-table3-halfyear-region-stats.csv")
selectivv[[4]] <- read_csv(file = "data/gig-table4-halfyear-activity-stats.csv")
```

BAEL data 
```{r}
bael <- read_csv("data/bael.csv") %>%
  na.omit() %>%
  unite(period, year:yearhalf, sep = ".") %>%
  rename(region_name = nazwa_miasta)
```

Figure 3: Number of monthly active users of selected apps in Poland by type and between 2018 and2020

```{r}
selectivv[[1]] %>%
  mutate(label = if_else(month == max(month), as.character(apps), NA_character_)) %>%
  ggplot(data = ., aes(x = month, y = number_of_users/1000, color = apps, group = apps)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(aes(label = label),
                  nudge_x = -0.25, 
                  direction = "y", 
                  hjust = "left",
                  box.padding = .1,
                  force = .8,
                  na.rm = TRUE,
                  show.legend = FALSE) + 
  scale_color_brewer(type = "qual", palette = "Dark2") + 
  labs(x = "Period", y = 'Number of users (thousands)', color = "App") + 
  facet_wrap(~category, scales = "free_y") +
  scale_x_date(limits = c(as.Date("2018-01-01"), as.Date("2021-09-01"))) + 
  theme(strip.text = element_text(size = 15)) ->  p1

p1

ggsave(plot = p1, file = "figures/fig-sel-month.png", width = 7, height = 4)
```

Figure 4 Number of monthly active users of selected apps in Poland by country, age and gender between 2018 and 2020

```{r}
selectivv[[2]] %>%
  count(category, apps, country, period, wt = n_unique_users) %>%  
  ggplot(data = ., aes(x = as.character(period), y = n/1000, group = country, color = country)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ category + apps) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Country") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2a

p2a

ggsave(plot = p2a, file = "figures/fig-sel-half-country.png", width = 8, height = 4)
```

```{r}
selectivv[[2]] %>%
  count(category, apps, age, period, wt = n_unique_users) %>%
  ggplot(data = ., aes(x = as.character(period), y = n/1000, group = age, color = age)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ category + apps) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Age group") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2b

p2b

ggsave(plot = p2b, file = "figures/fig-sel-half-age.png", width = 8, height = 4)

```

```{r}
selectivv[[2]] %>%
  count(category, apps, gender, period, wt = n_unique_users)  %>%
  ggplot(data = ., aes(x = as.character(period), y = n/1000, group = gender, color = gender)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ category + apps) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Gender") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2c

p2c

ggsave(plot = p2c, file = "figures/fig-sel-half-sex.png", width = 8, height = 4)

```

Figure 5 -- Point and 95\% interval of ratio of the number of active users and the LFS working 18--65 population by app for 2020HY1. Note that Y axis differs between the apps

```{r}
selectivv[[3]]  %>%
    filter(group == "Cities")  %>%
    mutate(period = as.character(period)) %>%
    select(category, period, region_name, apps, number_of_users) %>%
    left_join(
      bael %>%
        mutate(region_name = str_remove(region_name, "M. "))
    ) %>%
    filter(number_of_users > 0 ) -> cities_bael
```

```{r}
cities_bael %>%
  filter(region_name != "Sopot", period == "2020.1") %>%
  filter(!is.na(y))  %>%
  mutate(share = number_of_users/y*100,
         share_hi = number_of_users / (y - 1.96*rmse)*100,
         share_lo = number_of_users / (y + 1.96*rmse)*100) %>%
  mutate(apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                               "Takeaway", "Wolt", "Glover",  "Bolt Courier"))) %>%
  ggplot(data = ., aes(x = reorder(region_name, -y), y = share, ymin = share_lo, ymax = share_hi)) +
  geom_errorbar() +
  geom_point()  +
  facet_wrap(~apps, ncol = 4, scale = "free_y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 5))  +
  #scale_y_continuous(labels = scales::label_percent(accuracy = 2)) +
  labs(x = "City (ordered by the LFS working population size)", 
       y = "Number of users / LFS working population [%]") -> p3

p3

ggsave(plot = p3, file = "figures/fig-sel-cities-shares.png", width = 8, height = 4)

```

Figure 6 -- Average and standard deviation (Std. Dev.) of activity time by app type, app, weekdays and hours

```{r}
selectivv[[4]] %>%
  mutate(apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                             "Bolt Courier", "Takeaway", "Wolt", "Glover")),
         day = factor(day, c("Mondays-Thursdays", "Fridays-Sundays")))  %>%
  ggplot(data = ., aes(x = as.character(period), y = activity_time, color = apps, group = apps)) +
  geom_line() +
  geom_point() + 
  facet_nested(day ~ category +  hour + statistic)  +
  scale_color_brewer(type = "qual", palette = "Dark2", name = "Apps") +  
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(x = "Half-year", y = "Average time (in hours)") + 
  scale_y_continuous(breaks = seq(0,7,1))  -> p4

p4

ggsave(plot = p4, file = "figures/fig-sel-half-time-dev.png", width = 9, height = 5)

```

APPENDIX

Table 6 Number of active users in cities for 2020HY2

```{r}
selectivv[[3]] %>%
    filter(group == "Cities") %>%
    mutate(apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                                 "Takeaway", "Wolt", "Glover",  "Bolt Courier"))) %>%
    filter(number_of_users > 0, period == "2020.2")  %>% 
    select(period, apps, region_name, number_of_users) %>%
    spread(apps, number_of_users) %>% 
    select(-period) %>% 
    xtable(digits = 0) %>% 
    print.xtable(include.rownames = F, NA.string = "--")
```

Table 7 Direct estimates, its standard errors (in thousands) and sample sizes of the working population size (18--64) based on the Labour Force Survey between 2018HY1 and 2020HY

```{r}
bael %>%
  filter(period == "2018.1")  %>% 
  mutate(region_name = str_remove(region_name, "M. "),
         y = y / 1000,
         rmse = rmse / 1000,
         cv = rmse/y*100) %>%
  select(region_name, size = y, std = rmse, sample = n) %>%
  left_join(
    bael %>%
    filter(period == "2018.2")  %>% 
    mutate(region_name = str_remove(region_name, "M. "),
         y = y / 1000,
         rmse = rmse / 1000,
         cv = rmse/y*100) %>%
    select(region_name, size = y, std = rmse, sample = n),
    by = "region_name"
  ) %>%
  left_join(
    bael %>%
    filter(period == "2019.1")  %>% 
    mutate(region_name = str_remove(region_name, "M. "),
         y = y / 1000,
         rmse = rmse / 1000,
         cv = rmse/y*100) %>%
    select(region_name, size = y, std = rmse, sample = n),
    by = "region_name"
  )  %>%
  left_join(
    bael %>%
    filter(period == "2019.2")  %>% 
    mutate(region_name = str_remove(region_name, "M. "),
         y = y / 1000,
         rmse = rmse / 1000,
         cv = rmse/y*100) %>%
    select(region_name, size = y, std = rmse, sample = n),
    by = "region_name"
  )  %>%
  left_join(
    bael %>%
    filter(period == "2020.1")  %>% 
    mutate(region_name = str_remove(region_name, "M. "),
         y = y / 1000,
         rmse = rmse / 1000,
         cv = rmse/y*100) %>%
    select(region_name, size = y, std = rmse, sample = n),
    by = "region_name"
  )  %>%
  mutate_at(vars(starts_with("sample")), as.integer) %>%
  xtable(digits = 1) %>%
  print.xtable(include.rownames = F)

```